### 2016-03-05 paper/icfp16-submission/gallery.tex
misc

\begin{figure*}[ht]
\centering
\begin{minipage}{0.49\linewidth}
\centering
\begin{ecode}
let rec sumList xs = match xs with
  | []    -> []
  | y::ys -> y + __sumList ys__
\end{ecode}
% File "sumList.ml", line 3, characters 17-27:
\begin{verbatim}
Error: This expression has type
         'a list
       but an expression was expected of type
         int
\end{verbatim}
\end{minipage}
\begin{minipage}{0.49\linewidth}
\centering
\includegraphics[height=125px]{sumlist.png}
\end{minipage}
\caption{\texttt{sumList xs} should sum the elements of \texttt{xs}, but
  the student has returned \texttt{[]} in the base case instead of
  \texttt{0}. \ocaml deduces from the base case that \texttt{sumList}
  must return a \texttt{list} and thus blames the recursive call for
  producing a \texttt{list} instead of an \texttt{int}. \toolname's
  jump-compressed trace highlights the clash between the \texttt{[]} and
  the \texttt{+}, but also demonstrates in the third step that
  \texttt{sumList []} reduces to \texttt{[]}, which clarifies that the
  conflict is between the base case and recursive case.}
\label{fig:traces}
\end{figure*}

\begin{figure*}[ht]
\centering
\begin{minipage}{0.49\linewidth}
\centering
\begin{ecode}
let append xs ys = match xs with
  | []   -> ys
  | h::t -> h :: __t__ :: ys
\end{ecode}
% File "append.ml", line 3, characters 17-18:
\begin{verbatim}
Error: This expression has type
         'a list
       but an expression was expected of type
         'a
       The type variable 'a occurs inside 'a list
\end{verbatim}
\end{minipage}
\begin{minipage}{0.49\linewidth}
\centering
\includegraphics[height=75px]{append.png}
\end{minipage}
\caption{\texttt{append xs ys} should concatenate the two input lists,
  but the student has tried to cons the tail of \texttt{xs} onto
  \texttt{ys} instead of making a recursive call to \texttt{append}.
  \ocaml blames \texttt{t} for being a \texttt{'a list} instead of a
  \texttt{'a}, prompting an \emph{occurs-check} error. \toolname shows,
  more concretely, that \texttt{[] :: [2]} gets stuck because lists are
  homogeneous and \texttt{[]} is not an \texttt{int}.}
\label{fig:traces}
\end{figure*}

@@

let _ = wwhile (__f__, 2)
\end{ecode}
% File "wwhile.ml", line 10, characters 16-17:
\begin{verbatim}
Error: This expression has type
         int -> int * bool
       but an expression was expected of type
         'a * bool
\end{verbatim}
\end{minipage}
\begin{minipage}{0.49\linewidth}
\centering
\includegraphics[height=150px]{wwhile.png}
\end{minipage}
\caption{\texttt{wwhile (f, b)} should repeatedly apply the function
  \texttt{f} to the first element of its result pair, starting with the
  initial input \texttt{b}, until the second element is \texttt{false}.
  Unfortunately, the student has forgotten to apply \texttt{f} at all
  and just matches it directly against a pair. This faulty definition of
  \texttt{wwhile} still typechecks however, and \ocaml thus blames the
  \emph{call-site} on line 10. \toolname makes no assumptions about the
  input code and quickly draws the eye to the true error, the
  \texttt{match} expression, and highlights the conflict in matching a
  function against a pair pattern.}
\label{fig:traces}
\end{figure*}

>>>>

\begin{figure*}[ht]
\centering
\begin{minipage}{0.49\linewidth}
\centering
\begin{ecode}
let rec sumList xs = match xs with
  | []    -> []
  | y::ys -> y + __sumList ys__
\end{ecode}
% File "sumList.ml", line 3, characters 17-27:
\begin{verbatim}
Error: This expression has type
         'a list
       but an expression was expected of type
         int
\end{verbatim}
\end{minipage}
\begin{minipage}{0.49\linewidth}
\centering
\includegraphics[height=125px]{sumlist.png}
\end{minipage}
\caption{\texttt{sumList xs} should sum the elements of \texttt{xs}, but
  the student has returned \texttt{[]} in the base case instead of
  \texttt{0}. \ocaml deduces from the base case that \texttt{sumList}
  must return a \texttt{list} and thus blames the recursive call on line
  3 for producing a \texttt{list} instead of an  \texttt{int}.
  \toolname's jump-compressed trace highlights the clash
  between the \texttt{[]} and the \texttt{+}, but also demonstrates in
  the third step that \texttt{sumList []} reduces to \texttt{[]}, which
  clarifies that the conflict is between the base case and recursive
  case.}
\label{fig:traces}
\end{figure*}

\begin{figure*}[ht]
\centering
\begin{minipage}{0.49\linewidth}
\centering
\begin{ecode}
let append xs ys = match xs with
  | []   -> ys
  | h::t -> h :: __t__ :: ys
\end{ecode}
% File "append.ml", line 3, characters 17-18:
\begin{verbatim}
Error: This expression has type
         'a list
       but an expression was expected of type
         'a
       The type variable 'a occurs inside 'a list
\end{verbatim}
\end{minipage}
\begin{minipage}{0.49\linewidth}
\centering
\includegraphics[height=75px]{append.png}
\end{minipage}
\caption{\texttt{append xs ys} should concatenate the two input lists,
  but the student has tried to cons the tail of \texttt{xs} onto
  \texttt{ys} instead of making a recursive call to \texttt{append}.
  \ocaml blames \texttt{t} on line 3 for being a \texttt{'a list}
  instead of a \texttt{'a}, prompting an \emph{occurs-check}
  error. \toolname shows, more concretely, that \texttt{[] :: [2]} gets
  stuck because lists are homogeneous and \texttt{[]} is not an
  \texttt{int}.}
\label{fig:traces}
\end{figure*}

@@

let _ = wwhile (__f__, 2)
\end{ecode}
% File "wwhile.ml", line 10, characters 16-17:
\begin{verbatim}
Error: This expression has type
         int -> int * bool
       but an expression was expected of type
         'a * bool
\end{verbatim}
\end{minipage}
\begin{minipage}{0.49\linewidth}
\centering
\includegraphics[height=150px]{wwhile.png}
\end{minipage}
\caption{\texttt{wwhile (f, b)} should repeatedly apply the function
  \texttt{f} to the first element of its result pair, starting with the
  initial input \texttt{b}, until the second element is \texttt{false}.
  Unfortunately, the student has forgotten to apply \texttt{f} at all on
  line 2, and just matches it directly against a pair. This faulty
  definition of \texttt{wwhile} still typechecks however, and \ocaml
  thus blames the \emph{call-site} on line 10. \toolname makes no
  assumptions about the input code and quickly draws the eye to the true
  error, the \texttt{match} expression, and highlights the conflict in
  matching a function against a pair pattern.}
\label{fig:traces}
\end{figure*}
